The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba

### 2020-12-02 analysis/paper/paper.Rmd
updating paper

The family Gobiidae includes almost 2,000 species of marine, freshwater, and brackish fishes worldwide [@eschmeyer2020catalog]. This group is mostly represented by small-sized and bottom-dweller fishes found in tropical and subtropical waters. Such a small size along with cryptic body colors and the lack of diagnostic characters to distinguish among species and even genera is still a challenge for science in the 21st century. Only in the last ten years, more than a hundred gobiid species were described [@eschmeyer2020catalog]. However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematic of these fish taxa [@thacker2011phylogeny, @victor2014three, @keith2017schismatogobius].

Although most species in the family Gobiidae inhabit marine ecosystems, most species inhabiting freshwater or estuarine ecosystems display an amphidromous lifecycle [@myers1949usage, @bell1999overview]. This means that adults spawn in freshwater and when eggs hatch, they are taken downstream by river flow into the ocean. Then, after spending from a few weeks to a few months (depending on the species), postlarvae return to rivers. This moment of postlarval returning is particularly fascinating because million of them gather at river mouth as they start going upstream. This migratory behavior, particularly, when the postlarvae are going back to rivers, supports an artisanal fishery around the world [@bell1999overview, @mcdowall2007amphidromy]. Known as goby-fry fisheries, this is a common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean [@neal2009factors]. The problem with it is that during the last years insufficient management of these fisheries has led to considerable reduction in catches [@bell1999overview]. In addition, information of the identity of the species comprising these goby-fry fisheries in many places is still incomplete [@castellanos2011goby, @asis2013first].

@@

This paper was completely developed in RStudio, PBC (“RStudio”) [@rstudio2020]. R code and data are available as a research compendium for generating this paper using the following link: https://github.com/sheilarm23/rrcompendium.git.

@@

# checking that the tree is rooted
rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5) 
# getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
set.seed(123)
nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number

@@

```{r Sicydium_ML_tree, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*. Sequences highlited in green represent samples from Cuba. Numbers on branches represent boostrap values higher than 70.")}

Sicydium_ML_tree <-suppressWarnings(suppressMessages(print(ggtree(tree, size = .4) + 
  xlim(-0.02, 0.26) +
  geom_treescale(x=0, y=-1.5, offset = -4) + 
  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 1, col= "black") + 
  geom_cladelabel(node=216, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=108, label="Sicydium brevifile", color="blue", hjust = -0.01, vjust = 0.2, fontsize = 3.5, fontface="italic") + 
  geom_cladelabel(node=408, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=356, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=320, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=407, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
  geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
  geom_hilight(node=336, fill="green") +
  #geom_hilight(node=332, fill="green") +
  geom_hilight(node=135, fill="green") +
  geom_hilight(node=121, fill="green") +
  geom_hilight(node=120, fill="green") +
  geom_hilight(node=139, fill="green") +
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")))))

>>>>

The family Gobiidae includes almost 2,000 species of marine, freshwater, and brackish fishes worldwide [@eschmeyer2020catalog]. This group is mostly represented by small-sized and bottom-dweller fishes found in tropical and subtropical waters. Such a small size along with cryptic body colors and the lack of diagnostic characters to distinguish among species and even genera is still a challenge for science in the 21st century. Only in the last ten years, more than a hundred gobiid species were described [@eschmeyer2020catalog]. However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematic of these fish taxa [@thacker2011phylogeny; @victor2014three; @keith2017schismatogobius].

Although most species in the family Gobiidae inhabit marine ecosystems, most species inhabiting freshwater or estuarine ecosystems display an amphidromous lifecycle [@myers1949usage; @bell1999overview]. This means that adults spawn in freshwater and when eggs hatch, they are taken downstream by river flow into the ocean. Then, after spending from a few weeks to a few months (depending on the species), postlarvae return to rivers. This moment of postlarval returning is particularly fascinating because million of them gather at river mouth as they start going upstream. This migratory behavior, particularly, when the postlarvae are going back to rivers, supports an artisanal fishery around the world [@bell1999overview; @mcdowall2007amphidromy]. Known as goby-fry fisheries, this is a common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean [@neal2009factors]. The problem with it is that during the last years insufficient management of these fisheries has led to considerable reduction in catches [@bell1999overview]. In addition, information of the identity of the species comprising these goby-fry fisheries in many places is still incomplete [@castellanos2011goby; @asis2013first].

@@

This paper was completely developed in RStudio, PBC (“RStudio”) [@rstudio2020]. R code and data are available as a research compendium for generating this paper in the following link: https://github.com/sheilarm23/rrcompendium.git.

@@

# checking that the tree is rooted
rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5) 
# getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
set.seed(123)
nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number
nodeNumber

@@

```{r SicydiumTree, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*. Sequences highlited in green represent samples from Cuba. Numbers on branches represent boostrap values higher than 70.")}

ggtree(tree, size = .4) + 
  xlim(-0.02, 0.26) +
  geom_treescale(x=0, y=-1.5, offset = -4) + 
  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 1, col= "black") + 
  geom_cladelabel(node=215, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=108, label="Sicydium brevifile", color="blue", hjust = -0.01, vjust = 0.2, fontsize = 3.5, fontface="italic") + 
  geom_cladelabel(node=408, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=320, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=371, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=407, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
  geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
  # samples Sicydium_COI_Cuba2,4,6 were nested within this clde but the distance is too short to be displayed in the tree
  geom_hilight(node=191, fill="green") +
  geom_hilight(node=166, fill="green") +
  geom_hilight(node=177, fill="green") +
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))

# suppressWarnings(suppressMessages(print(Sicydium_ML_tree)))
